---
title: "Crime pre-processing"
author: "Abhiram Gaddam"
date: "2022-11-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
#library(readr)
library(lubridate)
library(ggridges)
```


1) Imputing null values 

For crime type (Abhi) - There are 5 values with missing "OFNS_DESC". Since they have valid PD_CD and PD_DESC, we can impute these values from other columns with the same PD_CD
      two values are for obscenity - 594 PD_CD. categorized as sex crimes
      one values for crime pos weap - 797 PD_CD. categorized as dangerous weapons 
      #OTHER
      one value for "place false bomb" - 648 PD_CD. No other examples of this crime. Ignore?
      one value for "noise" -  872 PD_CD. No other examples of this crime. Ignore?
      
```{r}
df<- read_csv('/Users/abhiramgaddam21/Desktop/EDAV/NYPD_Complaint_Data_Current__Year_To_Date_.csv')
#head(df)

df$OFNS_DESC[df$CMPLNT_NUM %in% c("248613125", "248290778")] <- "SEX CRIMES"
df$OFNS_DESC[df$CMPLNT_NUM %in% ("246605653")] <- "DANGEROUS WEAPONS"


df <- df[df$CMPLNT_NUM != "243170965",]
df <- df[df$CMPLNT_NUM != "245874611",]

```


2) Creating a new column to bucket crimes
```{r}

df <- df %>% 
  mutate(CRIME_CAT = case_when(
    OFNS_DESC %in% c("RAPE", "SEX CRIMES", "HARRASSMENT 2", "FELONY SEX CRIMES", "PROSTITUTION & RELATED OFFENSES") ~ "SEX CRIMES",
    OFNS_DESC %in% c("DANGEROUS DRUGS","CANNABIS RELATED OFFENSES","INTOXICATED & IMPAIRED DRIVING","ALCOHOLIC BEVERAGE CONTROL LAW","INTOXICATED/IMPAIRED DRIVING") ~ "DRUG AND ALCOHOL RELATED",
    OFNS_DESC %in% c("ROBBERY","GRAND LARCENY","THEFT-FRAUD","PETIT LARCENY","BURGLARY","GRAND LARCENY OF MOTOR VEHICLE","POSSESSION OF STOLEN PROPERTY","THEFT OF SERVICES","BURGLAR'S TOOLS","PETIT LARCENY OF MOTOR VEHICLE","OTHER OFFENSES RELATED TO THEF") ~ "THEFT OR BURGLARY",
    OFNS_DESC %in% c("DANGEROUS WEAPONS","MURDER & NON-NEGL. MANSLAUGHTER","KIDNAPPING & RELATED OFFENSES","HOMICIDE-NEGLIGENT,UNCLASSIFIE","HOMICIDE-NEGLIGENT-VEHICLE","KIDNAPPING","FELONY ASSAULT","ARSON","ASSAULT 3 & RELATED OFFENSES","UNLAWFUL POSS. WEAP. ON SCHOOL") ~ "MAJOR VIOLENT CRIMES",
    OFNS_DESC %in% c("CRIMINAL MISCHIEF & RELATED OF","UNAUTHORIZED USE OF A VEHICLE","FRAUDS","OFFENSES AGAINST PUBLIC SAFETY","DISORDERLY CONDUCT","JOSTLING","DISRUPTION OF A RELIGIOUS SERV","ESCAPE 3","OFF. AGNST PUB ORD SENSBLTY &","CRIMINAL TRESPASS","VEHICLE AND TRAFFIC LAWS","GAMBLING","OFFENSES AGAINST THE PERSON","OFFENSES INVOLVING FRAUD","FRAUDULENT ACCOSTING","ANTICIPATORY OFFENSES","LOITERING/GAMBLING (CARDS, DIC") ~ "FRAUD/GAMBLING AND MISC",
    OFNS_DESC %in% c("NYS LAWS-UNCLASSIFIED FELONY","MISCELLANEOUS PENAL LAW","FORGERY","OFFENSES AGAINST PUBLIC ADMINI","CHILD ABANDONMENT/NON SUPPORT","NYS LAWS-UNCLASSIFIED VIOLATION","OTHER STATE LAWS","OTHER STATE LAWS (NON PENAL LAW)","NEW YORK CITY HEALTH CODE","ADMINISTRATIVE CODE","OTHER STATE LAWS (NON PENAL LA","AGRICULTURE & MRKTS LAW-UNCLASSIFIED","ENDAN WELFARE INCOMP","OFFENSES RELATED TO CHILDREN") ~"OTHER"))

```


3) Basic counts and plots
```{r}
#head(df)

agg_tbl <- df %>% group_by(CRIME_CAT) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
df_counts <- agg_tbl %>% as.data.frame()
df_counts
View(df_counts)

#x <- c(4, 5, 6, 7, 8, 9, 10)
#counts <- c(78, 45, 50, 12, 30, 50)
#df <- data.frame(x=x, counts=counts)
#new_DF <- df[is.na(df$CRIME_CAT),]
#df_test <- read_csv('/Users/abhiramgaddam21/Desktop/EDAV/NYPD_Complaint_Data_Current__Year_To_Date_.csv')
#df_test <- df_test[is.na(df_test$CMPLNT_NUM),]
#View(df_test)
#plt <- ggplot(df_counts) + geom_bar(aes(x=CRIME_CAT, y=total_count), stat="identity")
#print(plt)
```







