---
title: "Data Preparation"
author: "Group32"
date: "2022-11-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo= TRUE, warning = TRUE, message = TRUE)
#knitr::opts_chunk$set(echo= TRUE, warning = FALSE, message = FALSE)
```

```{r libraries}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggridges)
library(geosphere)

#Unused Libraries:
#library(readr)
```

```{r DataLoading}
#Group32 - This file has been added to gitignore so it will not be uploaded. So we are on the same page and can run the same code, add the csv to your local project folder in a new subfolder "data". 
df_raw <- read_csv('./data/NYPD_Complaint_Data_Current__Year_To_Date_.csv')
#View(df_raw)
#head(df_raw)

```

#Data Standardization Contents

1) Re-categorize and clean Offense Type
2) Derive Date/Time fields
3) 



#Offense Type

##Imputing null values 

For crime type (Abhi) - There are 5 values with missing "OFNS_DESC". Since they have valid PD_CD and PD_DESC, we can impute these values from other columns with the same PD_CD
      two values are for obscenity - 594 PD_CD. categorized as sex crimes
      one values for crime pos weap - 797 PD_CD. categorized as dangerous weapons 
      one value for "place false bomb" - 648 PD_CD. No other examples of this crime. Ignore?
      one value for "noise" -  872 PD_CD. No other examples of this crime. Ignore?
      
```{r DataPrep1_CrimeType_Nulls}
df <- df_raw

df$OFNS_DESC[df$CMPLNT_NUM %in% c("248613125", "248290778")] <- "SEX CRIMES"
df$OFNS_DESC[df$CMPLNT_NUM %in% ("246605653")] <- "DANGEROUS WEAPONS"

```

##Creating a new column to bucket crimes

```{r DataPrep1_CrimeType_Buckets}
df <- df %>% 
  mutate(CRIME_CAT = case_when(
    OFNS_DESC %in% c("RAPE", "SEX CRIMES", "HARRASSMENT 2", "FELONY SEX CRIMES", "PROSTITUTION & RELATED OFFENSES") ~ "SEX CRIMES",
    OFNS_DESC %in% c("DANGEROUS DRUGS","CANNABIS RELATED OFFENSES","INTOXICATED & IMPAIRED DRIVING","ALCOHOLIC BEVERAGE CONTROL LAW","INTOXICATED/IMPAIRED DRIVING") ~ "DRUG AND ALCOHOL RELATED",
    OFNS_DESC %in% c("ROBBERY","GRAND LARCENY","THEFT-FRAUD","PETIT LARCENY","BURGLARY","GRAND LARCENY OF MOTOR VEHICLE","POSSESSION OF STOLEN PROPERTY","THEFT OF SERVICES","BURGLAR'S TOOLS","PETIT LARCENY OF MOTOR VEHICLE","OTHER OFFENSES RELATED TO THEF") ~ "THEFT OR BURGLARY",
    OFNS_DESC %in% c("DANGEROUS WEAPONS","MURDER & NON-NEGL. MANSLAUGHTER","KIDNAPPING & RELATED OFFENSES","HOMICIDE-NEGLIGENT,UNCLASSIFIE","HOMICIDE-NEGLIGENT-VEHICLE","KIDNAPPING","FELONY ASSAULT","ARSON","ASSAULT 3 & RELATED OFFENSES","UNLAWFUL POSS. WEAP. ON SCHOOL","MURDER & NON-NEGL. MANSLAUGHTER") ~ "MAJOR VIOLENT CRIMES",
    OFNS_DESC %in% c("CRIMINAL MISCHIEF & RELATED OF","UNAUTHORIZED USE OF A VEHICLE","FRAUDS","OFFENSES AGAINST PUBLIC SAFETY","DISORDERLY CONDUCT","JOSTLING","DISRUPTION OF A RELIGIOUS SERV","ESCAPE 3","OFF. AGNST PUB ORD SENSBLTY &","CRIMINAL TRESPASS","VEHICLE AND TRAFFIC LAWS","GAMBLING","OFFENSES AGAINST THE PERSON","OFFENSES INVOLVING FRAUD","FRAUDULENT ACCOSTING","ANTICIPATORY OFFENSES","LOITERING/GAMBLING (CARDS, DIC") ~ "FRAUD/GAMBLING AND MISC",
    OFNS_DESC %in% c("NYS LAWS-UNCLASSIFIED FELONY","MISCELLANEOUS PENAL LAW","FORGERY","OFFENSES AGAINST PUBLIC ADMINI","CHILD ABANDONMENT/NON SUPPORT","NYS LAWS-UNCLASSIFIED VIOLATION","OTHER STATE LAWS","OTHER STATE LAWS (NON PENAL LAW)","NEW YORK CITY HEALTH CODE","ADMINISTRATIVE CODE","OTHER STATE LAWS (NON PENAL LA","AGRICULTURE & MRKTS LAW-UNCLASSIFIED","ENDAN WELFARE INCOMP","OFFENSES RELATED TO CHILDREN") ~"OTHER"))

df$CRIME_CAT[df$CMPLNT_NUM %in% c("243170965", "245874611")] <- "OTHER"

#No Nulls
#View(df[is.na(df$CRIME_CAT),])

```




#Date/Time

##Cleaning null values/ranges
Per the data dictionary, there is a "from" date and a "to" when the exact time is unknown. There is also a "report date" for when the crime was reported. Using these fields, we can derive a new clean field that is the assumed date of the incident. Assumptions: When there is a range, we will use the "from" date because is populated well and will on average approximate the frequency of crime over time; when from date is null, we will use the report date. 

```{r DataPrep2_DateTimeFields}

df <- df %>%
      mutate(
         #use from date and report date if null. If to_date then just use from date and we can argue it averages out since new reports will start as other end
          Incident_Date_raw = case_when (is.null(CMPLNT_FR_DT) ~ RPT_DT
                                     ,CMPLNT_FR_DT == "(null)" ~ RPT_DT
                                     ,TRUE  ~ CMPLNT_FR_DT
                                    )
         #flag if estimated (ie from date is null or to date is populated)
         ,Incident_Date_Estimated_Flag = case_when ( is.null(CMPLNT_FR_DT) ~ 'Y'
                                                     ,CMPLNT_FR_DT == "(null)" ~ 'Y'
                                                     ,!is.null(CMPLNT_TO_DT) ~ 'Y'
                                                     ,CMPLNT_TO_DT != "(null)" ~ 'Y'
                                                     ,TRUE ~ 'N'
                                                   )
  ) 

#Convert to times
df <- df %>% 
      mutate(
        #creating date and time together for lubridate
         Incident_Date = as.Date(Incident_Date_raw, format = '%m/%d/%Y')
        ,Incident_Datetime = as.POSIXct(paste(Incident_Date_raw,CMPLNT_FR_TM), format = '%m/%d/%Y %H:%M:%S')
      ) %>%
      mutate(
         Incident_HourTime = hour(Incident_Datetime) + minute(Incident_Datetime)/60
        ,Incident_Month = month(Incident_Date)
        ,Incident_DayOfWeek = wday(Incident_Date, label = TRUE, abbr = TRUE)
      )

```

#GeoLocation Fields

```{r DataPrep3_DistanceFields}
#add column for Columbia location
CU_Latitude = 40.807384
CU_Longitude = -73.963036

#df_test = df[1:10,]
#df_test$dist_to_CU <- apply(df_test, 1, function(x)distm(c(x[which( colnames(df)=="Longitude")],x[which(colnames(df)=="Latitude")]),c(x[which( colnames(df)=="CU_Longitude")],x[which( colnames(df)=="CU_Latitude")]),fun = distGeo))

df$dist_to_CU <- apply(df, 1, function(x)distm(
   c(x[which( colnames(df)=="Longitude")],x[which(colnames(df)=="Latitude")])
  ,c(CU_Longitude,CU_Latitude)
  ,fun = distGeo)
  )

```
--#TODO can we speed this up? Takes about 5 mins to run



#Other Derivations

Here we will add additional fields for more classification
--#TODO add more description

```{r DataPrep4_MoreFields}

df$LOC_OF_OCCUR_DESC[df$LOC_OF_OCCUR_DESC=="(null)"]<-NA

df <- df %>% 
      mutate(
         #get a flag for outside vs inside
         Inside_Outside = case_when ( 
                   LOC_OF_OCCUR_DESC %in% c("FRONT OF" , "OPPOSITE OF" , "REAR OF") ~ "OUTSIDE"
                  ,LOC_OF_OCCUR_DESC %in% c("INSIDE") ~ "INSIDE"
                  #,LOC_OF_OCCUR_DESC == "(null)" ~ NULL #TODO -cant figure this out error
                  ,TRUE ~ LOC_OF_OCCUR_DESC  
                  )
         #if victim was a person (not a business/govt)
         ,VIC_Individual_Flag = case_when (
                  VIC_SEX %in% c("M","F","L") ~ 'Y'
                  ,TRUE ~ 'N'
                  )
      ) %>% 
      mutate(
         Complaint_Count = 1     #maybe want to add like a intensity value or something?
#TODO, make this field a little better/check values?
         ,Premise_Derived = case_when (
           Inside_Outside == 'INSIDE' ~ 'INSIDE'
           ,PREM_TYP_DESC =="RESIDENCE - APT. HOUSE" & (Inside_Outside == "(null)" | is.null(Inside_Outside)) ~ 'INSIDE'
           ,!is.null(PARKS_NM) & PARKS_NM != "(null)" ~ 'PARK'
           ,PREM_TYP_DESC %in% c("TRANSIT - NYC SUBWAY","BUS (NYC TRANSIT)","TRANSIT FACILITY (OTHER)") ~ 'SUBWAY'
           ,TRUE ~ 'STREET'
         )
      )

```


#Select and Rename Columns in a Usuable Table

--#TODO add more description

```{r DataPrep5_Renaming}
#Not required
```


#Filter Table for Relevance

--#TODO add more description
Date - We only have crimes that took place (as defined by derived field above) in 1/1/2022 onward (this should remove 5%) (Devan)
  Crime Type - Use the derived field from above 
  Location? - Tentative, only look within 3.2km of Columbia (per derived field above)
  
```{r DataPrep6_Filter}
df_filter <- df %>% filter(  Incident_Date >= as.Date('2022/01/01')  ) %>%
  filter (
    !CRIME_CAT %in% c("OTHER")
  )

#TODO- distance filter?

#View(df_filter)
#unique(df_filter$CRIME_CAT)

```



